---
title: "run_analysis.R"
author: "Mario Vigil"
date: "8/29/2020"
output: html_document
---

```{r}
library(tidyverse)
library(codebook)
library(future)
```



Dowload data / set directory
```{r}
if(!file.exists("~/Desktop/Coursera/CleaningData")){dir.create("~/Desktop/Coursera/CleaningData")}
Url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(Url,destfile="~/Desktop/Coursera/CleaningData/Dataset.zip",method="curl")
unzip(zipfile="~/Desktop/Coursera/CleaningData/Dataset.zip",exdir="~/Desktop/Coursera/CleaningData")
path <- file.path("~/Desktop/Coursera/CleaningData" , "UCI HAR Dataset")
files<-list.files(path, recursive=TRUE)
files
```

# 1. Merges the training and the test sets to create one data set.

Data
```{r}
# Reading data and column names
features <- read.table(file.path(path, "features.txt"), 
                       col.names = c("n","functions")) # N and the different functions

activities <- read.table(file.path(path, "activity_labels.txt"), 
                         col.names = c("code", "activity")) # Code and value for each activity

YTest  <- read.table(file.path(path, "test" , "Y_test.txt" ),
                     col.names = "code") # Test labels

YTrain <- read.table(file.path(path, "train", "Y_train.txt"),
                     col.names = "code") # Training labels

XTest  <- read.table(file.path(path, "test" , "X_test.txt" ),
                     col.names = features$functions) # Test values

XTrain <- read.table(file.path(path, "train", "X_train.txt"),
                     col.names = features$functions) # Train values 

SubjectTrain <- read.table(file.path(path, "train", "subject_train.txt"),
                           col.names = "subject") # subjects in train set

SubjectTest  <- read.table(file.path(path, "test" , "subject_test.txt"),
                           col.names = "subject") # subjects in test set

```

Merging
```{r}
# Row Bind
subject <- rbind(SubjectTrain, SubjectTest)
YData<- rbind(YTest, YTrain)
XData<- rbind(XTest, XTrain)

# ColumnBind
MergedData <- cbind(subject, YData, XData)
head(MergedData)
```


# 2. Extract mean and standard deviation

```{r}
MeanSD <- MergedData %>% 
        select(subject, code, contains("mean"), contains("std"))
```


# 3. Uses descriptive activity names to name the activities in the data set
```{r}

MeanSD$code <- activities[MeanSD$code, 2]
```

# 4. Uses descriptive activity names to name the activities in the data set

```{r}
names(MeanSD)[2] = "activity"
names(MeanSD)<-gsub("^t", "time", names(MeanSD))
names(MeanSD)<-gsub("^f", "frequency", names(MeanSD))
names(MeanSD)<-gsub("Acc", "Accelerometer", names(MeanSD))
names(MeanSD)<-gsub("Gyro", "Gyroscope", names(MeanSD))
names(MeanSD)<-gsub("Mag", "Magnitude", names(MeanSD))
names(MeanSD)<-gsub("BodyBody", "Body", names(MeanSD))

names(MeanSD)
```

# 4. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}

AcceleramatorClean <- MeanSD %>%
    group_by(subject, activity) %>%
    summarise_all(funs(mean))
write.table(AcceleramatorClean, "AcceleramatorClean.txt", row.name=FALSE)
```

# Code book

```{r}
book <- codebook(AcceleramatorClean)

book
```







