---
title: "run_analysis.R"
author: "Mario Vigil"
date: "8/29/2020"
output: html_document
---

```{r}
library(tidyverse)
```



Dowload data / set directory
```{r}
if(!file.exists("~/Desktop/Coursera/CleaningData")){dir.create("~/Desktop/Coursera/CleaningData")}
Url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(Url,destfile="~/Desktop/Coursera/CleaningData/Dataset.zip",method="curl")
unzip(zipfile="~/Desktop/Coursera/CleaningData/Dataset.zip",exdir="~/Desktop/Coursera/CleaningData")
path <- file.path("~/Desktop/Coursera/CleaningData" , "UCI HAR Dataset")
files<-list.files(path, recursive=TRUE)
files
```

# 1. Merges the training and the test sets to create one data set.

Data
```{r}
ActivityTest  <- read.table(file.path(path, "test" , "Y_test.txt" ),header = FALSE)
ActivityTrain <- read.table(file.path(path, "train", "Y_train.txt"),header = FALSE)
SubjectTrain <- read.table(file.path(path, "train", "subject_train.txt"),header = FALSE)
SubjectTest  <- read.table(file.path(path, "test" , "subject_test.txt"),header = FALSE)
FeaturesTest  <- read.table(file.path(path, "test" , "X_test.txt" ),header = FALSE)
FeaturesTrain <- read.table(file.path(path, "train", "X_train.txt"),header = FALSE)

```

Merging
```{r}
# Row Bind
subject <- rbind(SubjectTrain, SubjectTest)
activity<- rbind(ActivityTrain, ActivityTest)
features<- rbind(FeaturesTrain, FeaturesTest)
```

Variable names 
```{r}
names(subject)<-"subject"
names(activity)<- "activity"
FeatureRows <- read.table(file.path(path, "features.txt"),head=FALSE)
names(features)<- FeatureRows$V2
```

```{r}
SA <- cbind(subject, activity)
total <- cbind(features, SA)
```

# 2. Extract mean and standard deviation

```{r}
MeanSD <- grep('mean|std', features)
MeanSD <- total[,c(1,2,MeanSD + 2)]
```

# 3. Uses descriptive activity names to name the activities in the data set
```{r}
ActivityNames <- read.table("~/Desktop/Coursera/CleaningData/UCI HAR Dataset/activity_labels.txt", header = FALSE)
ActivityNames
```

# 4. Uses descriptive activity names to name the activities in the data set

```{r}
names(total)<-gsub("^t", "time", names(total))
names(total)<-gsub("^f", "frequency", names(total))
names(total)<-gsub("Acc", "Accelerometer", names(total))
names(total)<-gsub("Gyro", "Gyroscope", names(total))
names(total)<-gsub("Mag", "Magnitude", names(total))
names(total)<-gsub("BodyBody", "Body", names(total))

names(total)
```

# 4. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}

total<-aggregate(. ~subject + activity, total, mean)
total<-total[order(total$subject,total$activity),]
write.table(total, file = "total.txt",row.name=FALSE)
```






